---
title: "frost2thredds.Rd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(esd)
```

## R Markdown

This is an R Markdown containing R utilities in order to read weather station data from FROST and save it to a NetCDF4. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

The R utilities have been extracted from the esd R package (https://esd.met.no)

```{r define.elements}
metno.frost.ele <- function() {
  # TODO: wind direction is here so we can get DD06, DD12 and DD18 and perform an average.. how?
  x <- rbind(
    c("601" , "Precipitation"		, "1"	, "mm"		, "sum(precipitation_amount *)"),
    c("401" , "Sea level pressure"	, "1"	, "hPa"		, "mean(surface_air_pressure *)"),
    c("402" , "Sea level pressure"	, "1"	, "hPa"		, "min(surface_air_pressure *)"),
    c("403" , "Sea level pressure"	, "1"	, "hPa"		, "max(surface_air_pressure *)"),
    c("901" , "Snow depth"		, "1"	, "cm"		, "surface_snow_thickness"),
    c("101" , "Mean temperature"	, "1"	, "degree*C"	, "mean(air_temperature *)"),
    c("111" , "Maximum temperature"	, "1"	, "degree*C"	, "max(air_temperature *)"),
    c("121" , "Minimum temperature"	, "1"	, "degree*C"	, "min(air_temperature *)"),
    c("501" , "Wind speed"		, "1"	, "m/s"		, "mean(wind_speed *)"),
    c("502" , "Wind direction"		, "1"	, "degrees"	, "wind_from_direction"),
    c("503" , "Wind Gust"		, "1"	, "m/s"		, "max(wind_speed_of_gust *)")
  )
  y <- data.frame(element=x[,1] , longname = x[,2] , scale_factor = x[,3] , unit = x[,4] , param = x[,5] , source = "METNO.FROST" , stringsAsFactors = FALSE)
  return(y)
}

meta <- metno.frost.ele()
```

```{r read.elements}
#' Download METNO station metadata using frost.met.no
#' 
#' Where there are multiple measuring periods registered for the parameter,
#' only the earliest start time and the latest end time are used.
#' 
#' @aliases metno.frost.meta.day metno.frost.meta.month
#' 
#' @param param Vector of parameters
#' @param save2file if TRUE, save metadata in a local file
#' @param verbose if TRUE, print diagnostics
#' @param \dots additional arguments  
#' @param url The URL to the webpage to request new client credentials at Frost API
#'
#' @return A meta data matrix object for all stations in METNO's collection
#' that have measured any of the given parameters. Start and end time are included. 
#'
#' @author K. Tunheim
#'
#' @keywords parameter,metadata,metno,norway,frost
#'
#' @examples
#' # Fetch all stations' measuring periods of the t2m parameter
#' metno.frost.meta.min(param=c('t2m'))
#' # Fetch all stations' measuring periods of all available parameters
#' metno.frost.meta.min()

metno.frost.meta.default <- function(keyfile='~/.FrostAPI.key', param=c("t2m"), 
                                     timeresolutions="P1M", levels="default", timeoffsets="default", 
                                     performancecategories="A,B,C", exposurecategories="1,2", 
                                     url='https://frost.met.no/auth/requestCredentials.html',
                                     browser="firefox", verbose = FALSE) {
  if(verbose) print("metno.frost.meta.default")
  if (!requireNamespace("jsonlite", quietly = TRUE)) {
    stop("Package 'jsonlite' needed to use 'meta.frost.meta.default'. Please install it.")
  } else {
    
    # KMP 2020-01-22: enable timeresolutions notation monthly and daily
    timeresolutions <- switch(toupper(timeresolutions), 
                              "MONTHLY"="P1M", "MONTH"="P1M",
                              "DAILY"="P1D", "DAY"="P1D",
                              "MINUTE"="PT1M", "MIN"="PT1M",
                              timeresolutions)
    
    # convert all param to local param names
    getparam1 <- function(x) {
      withstar <- ele2param(x, src="METNO.FROST")$param
      gsub('*', timeresolutions, withstar, fixed=TRUE)
    }
    ele <- sapply(param, esd2ele)
    param1s <- sapply(ele, getparam1)
    names(param1s) <- ele
    strparam <- paste0(param1s, collapse=",")
    if (verbose) print(strparam)
    
    # Get a client_id
    if (file.exists(keyfile)) {
      if (verbose) print(paste('Read client ID from',keyfile))
      frostID <- readLines(keyfile) 
    } else { 
      if (verbose) print(paste('Generate new client ID from',url))  
      system(paste(browser,url))
      frostID <- rep("",2)
      frostID[1] <- readline('Please give me the first key:')
      frostID[2] <- readline('Please give me the second key:')
      writeLines(frostID,con=keyfile)
    }
    
    url1 <- paste0(
      "https://", 
      frostID[1],
      #client_id, 
      "@frost.met.no/",
      "sources/v0.jsonld",
      "?types=SensorSystem",
      "&country=Norge",
      "&validtime=0000-01-01/9999-01-01",
      "&fields=id,name,masl,country,county,countyId,municipality,municipalityId,geometry"
    )
    # KT 2020-05-25 - fetch all stations in and around Svalbard too
    url_sj <- paste0(
      "https://",
      frostID[1],
      "@frost.met.no/",
      "sources/v0.jsonld",
      "?types=SensorSystem",
      "&country=Svalbard og Jan Mayen",
      "&validtime=0000-01-01/9999-01-01",
      "&fields=id,name,masl,country,county,countyId,municipality,municipalityId,geometry"
    )
    url2 <- paste0(
      "https://",
      frostID[1],
      "@frost.met.no/",
      "observations/availableTimeSeries/v0.jsonld",
      "?elements=", strparam,
      "&timeresolutions=", timeresolutions,
      "&levels=", levels,
      "&timeoffsets=", timeoffsets,
      "&performancecategories=", performancecategories,
      "&exposurecategories=", exposurecategories,
      "&fields=sourceId,elementId,validFrom,validTo"
    )
    if (verbose) {
      print(url1)
      print(url_sj)
      print(url2)
    }
    
    # KT 2020-05-26: getting data from both Norge and Svalbard and Jan Mayen
    #browser()
    xs_no <- jsonlite::fromJSON(URLencode(url1), flatten=TRUE)
    xs_sj <- jsonlite::fromJSON(URLencode(url_sj), flatten=TRUE)
    xs1 <- rbind(xs_no$data, xs_sj$data)
    xs1$lon = sapply(xs1$geometry.coordinates, function(x) x[1])
    xs1$lon[sapply(xs1$lon, is.null)] <- NA
    xs1$lon <- unlist(xs1$lon)
    xs1$lat = sapply(xs1$geometry.coordinates, function(x) x[2])
    xs1$lat[sapply(xs1$lat, is.null)] <- NA
    xs1$lat <- unlist(xs1$lat)
    df1 <- xs1[c("id","name","country","lon","lat","masl","municipality","municipalityId","county","countyId")]
    
    xs2 <- jsonlite::fromJSON(URLencode(url2), flatten=TRUE)
    df2 <- xs2$data
    df2$sourceId = substring(df2$sourceId, 1, nchar(df2$sourceId)-2)
    df <- data.frame(NULL)
    for (i in 1:length(param1s)) {
      # KT 2020-05-26: preserve NA as latest validTo
      dfparam = df2[df2$elementId == param1s[i], ]
      dfparam$validTo[is.na(dfparam$validTo)] = "9999-12-31T00:00:00.000Z"
      validFrom = try(aggregate(validFrom ~ sourceId, data=dfparam, min), silent=TRUE)
      validTo = try(aggregate(validTo ~ sourceId, data=dfparam, max), silent=TRUE)
      validTo[validTo=="9999-12-31T00:00:00.000Z"] <- NA
      if (class(validFrom) != "try-error" & length(validFrom) > 0) {
        validFrom$validFrom <- as.Date(validFrom$validFrom)
        validTo$validTo <- as.Date(validTo$validTo)
        
        period = merge(validFrom, validTo, by='sourceId', all.x=TRUE)
        stperiod = merge(df1, period, by.x="id", by.y="sourceId")
        
        colnames(stperiod) = c("station_id","location","country","lon","lat","altitude",
                               "municipality","municipalityid","county","countyid","start","end")
        
        stperiod$element <- rep(names(param1s[i]),length(stperiod$station_id))
        
        # convert to UTM
        utmZone <- 33
        XY <- LatLon2UTM(lat=stperiod$lat, lon=stperiod$lon, zone=utmZone)
        stperiod$utm_east  <- XY[[1]]
        stperiod$utm_north <- XY[[2]]
        stperiod$utm_zone  <- rep(utmZone, length(stperiod$station_id))
        
        df <- rbind(stperiod, df, stringsAsFactors=FALSE)
      }
    }
    #invisible(df)
    
    ## Same format as station.meta
    var <- df$element
    for(element in unique(df$element)) {
      var[df$element==element] <- esd2ele(element)
    }
    cntr <- sapply(df$country, function(x) switch(x, "Norge"="NORWAY", x))
    X <- data.frame("station_id"=gsub("[A-Z]|[a-z]","",df$station_id),
                    "location"=df$location,
                    "country"=cntr,
                    "longitude"=df$lon,
                    "latitude"=df$lat,
                    "altitude"=df$altitude,
                    "element"=df$element,
                    "start"=strftime(df$start, format="%Y"),
                    "end"=strftime(df$end, format="%Y"),
                    "source"=switch(timeresolutions,
                                    "P1D"="METNOD.FROST",
                                    "P1M"="METNOM.FROST",
                                    "PT1M"="METNO1M.FROST"),
                    "wmo"=rep(NA,length(df$station_id)),
                    "quality"=rep(NA,length(df$station_id)),
                    "variable"=var, stringsAsFactors=FALSE)
    attr(X,"metnoURLs") <- "http://frost.met.no"
    attr(X,"author") <- "K. Tunheim & K. Parding"
    attr(X,"date") <- Sys.time()
    attr(X,"history") <- history.stamp(X)
    class(X) <- c("stationmeta", class(X))
    invisible(X)
  }
}

# Do not export - not in use
metno.frost.meta.minute <- function(param=c("t2m","precip","tmin","tmax","slp","pon","pox","fg","fx"),                                     save2file=FALSE, path=NULL, verbose=FALSE, ...) {
  
  if(verbose) print("metno.frost.meta.min")
  X <- metno.frost.meta.default(param=param, timeresolutions="PT1M", verbose=verbose, ...)
  filename <- "meta.metno.frost.1min.rda"
  attr(X, "source") <- "METNO.FROST.MINUTE"
  attr(X, "version") <- NA
  attr(X, "URL") <- "http://frost.met.no"
  attr(X, "file") <- filename
  attr(X, "cite") <- ""
  attr(X, "date") <- date()
  attr(X,"call") <- match.call()
  attr(X, "history") <- history.stamp(X)
  if (save2file) {
    meta.metno.frost.min <- X
    if(!is.null(path)) filename <- file.path(path,filename)
    save(meta.metno.frost.min, file=filename)
    rm("meta.metno.frost.min")
  }
  invisible(X)
}

#meta <- metno.frost.meta.minute(param='precip',save2file = FALSE,verbose = TRUE)
```

## saving to NetCDF files

You can also embed plots, for example:

```{r historical, echo=FALSE}


# Create historical files

#it <- as.POSIXct(c("2019-02-18 00:00:00","2021-01-31 00:05:00"))


meta <- metno.frost.meta.minute(param = 'precip', save2file=FALSE, verbose=TRUE)

generate.1 <- function(param = 'precip', it ,FUN='sum',verbose = TRUE) {
  #it <- seq(it, it, by = 5 * 60)
  x <- station(x = meta, param = param, it = it, src='METNO1M.FROST',verbose = TRUE)
  #browser()
  if (!is.null(FUN)) {
    if (verbose) print(paste('Apply ',FUN,' over time'))
    x <- map.station(x, FUN = FUN, plot = FALSE, verbose = verbose)
    index(x) <- as.POSIXct(attr(x,'period'))[2] + 60
    }
  
  # it <- gsub("[[:punct:]]", "", it)
  # it <- gsub(' ','T',it)
  
  # save as netcdf
  frost2ncdf4(x,verbose = TRUE)
}

generate.all <- function(){
  it.start <- as.POSIXct("2019-01-01 00:00:00")
  it.end <- as.POSIXct(Sys.time())
  
  it <- seq(from = it.start,
            to   = it.end,
            by   = 5*60) # 5 minutes interval
  
  # Convert to days
  vdays <- seq(from = round(it.start,units='days'),
               to   = round(it.end,units='days') ,
               by   ='day')
  
  for (i in 1835:(length(it)-1)) {
    it.i <- c(it[i],it[i+1])
    generate.1(it = it.i)
  }
}
```

## Realtime reading and saving

```{r realtime, echo=FALSE}
it <- seq(from = as.POSIXct("2019-01-01 00:00:00"),
          to   = as.POSIXct(Sys.time()),
          by   = 5*60) # 5 minutes

# Create historical files

it <- as.POSIXct(c("2021-02-18 00:00:00","2021-02-18 00:05:00"))
x <- station(x = meta, param = 'precip', it = it, src='METNO1M.FROST',verbose = TRUE)
gsub("[[:punct:]]", "", it[2])
write2ncdf4(x,file = 'precipitation-mm-norway.frost2ncdf.20210219T091500Z.nc')
```

## Write to NetCDF4 file
```{r write2ncdf}
frost2ncdf4 <- function(x, FUN= 'sum', namelength=24, prec='short', missval = -999, 
                        verbose = TRUE, force = TRUE) {
  
  if (verbose) print('Define dimensions')
  if (verbose) print(stid(x))
  dimS <- ncdim_def(name     = "id", 
                    units    = "",  
                    vals     = 1:length(stid(x)),
                    longname = "increment station number",
                    unlim    = TRUE)
  dimT <- ncdim_def(name     = "time",
                    longname = "time",
                    units    = "seconds since 1970-01-01 00:00:00 +00:00", 
                    vals     = as.numeric(difftime(index(x), 
                                                   as.POSIXct('1970-01-01 00:00:00',
                                                              format= "%Y-%m-%d %H:%M:%OS"), 
                                                   units = 'secs')), 
                    calendar = 'standard',
                    unlim    = TRUE)
  
  dimnchar <-ncdim_def(name  = "nchar",
                       units = "", 
                       vals  = 1:namelength,
                       create_dimvar = FALSE )
  # Generate path and filename 
  stamp <- gsub(pattern     = "[[:punct:]]", 
                replacement = "", 
                x           = paste0(gsub(' ','T',as.character(index(x))),'Z')) # e.g. 20210219T091500Z
  
   # Path to file
  path <- file.path(attr(x,'longname'),
                    as.character(year(x)), # Year subfolder
                    ifelse(nchar(as.character(month(x))) == 1,paste0('0',month(x)),month(x)),
                    ifelse(nchar(as.character(day(x))) == 1,paste0('0',day(x)),day(x))) # Month subfolder
  if (!file.exists(path))
    dir.create(path = path, recursive = TRUE)
  
  filename <-  paste0(unique(attr(x,'longname')),'-',
                      unique(unit(x)),'-',
                      unique(cntr(x)),
                      '.frost2ncdf.', 
                      stamp,'.nc')
  
  if (verbose) {
    print('Define variable')
    print(paste('create netCDF-file',filename))
  }
  
  if (verbose) 
    print('Define the netCDF structure')
  
  latid <- ncvar_def(name     = "lat",
                     dim      = list(dimS), 
                     units    = "degrees_north", 
                     missval  = missval,
                     longname = "latitude", 
                     prec     = "float",
                     verbose  = verbose)
  
  lonid <- ncvar_def(name     = "lon",
                     dim      = list(dimS), 
                     units    = "degrees_east", 
                     missval  = missval,
                     longname = "longitude", 
                     prec     = "float",
                     verbose  = verbose)
  
  xid <- ncvar_def(name     = "X",
                   dim      = list(dimS), 
                   units    = "meters", 
                   missval  = missval,
                   longname = "X coordinate to the east ", 
                   prec     = "float",
                   verbose  = verbose)
  
  yid <- ncvar_def(name     = "Y",
                   dim      = list(dimS), 
                   units    = "meters", 
                   missval  = missval,
                   longname = "Y coordinate to the North", 
                   prec     = "float",
                   verbose  = verbose)
  
  altid <- ncvar_def(name     = "alt",
                     dim      = list(dimS), 
                     units    = "meters", 
                     missval  = missval,
                     longname = "altitude", 
                     prec     = prec,
                     verbose  = verbose)
  
  locid <- ncvar_def(name     = "loc",
                     dim      = list(dimnchar,dimS),
                     units    = "",
                     prec     = "char",
                     longname = "location name",
                     verbose  = verbose)
  
  stid <- ncvar_def(name      = "stid", 
                    dim       = list(dimnchar,dimS),
                    units     = "",
                    prec      = "char",
                    longname  = "station identification number",
                    verbose   = verbose)
  
  cntrid <- ncvar_def(name    = "cntr",
                      dim     = list(dimnchar,dimS),
                      units   = "",
                      prec    = "char",
                      longname= "country name",
                      verbose = verbose)
  
  if (verbose) {
    print(paste('ncvar:',varid(x)[1]))
    print(unit(x))
  }
  
  
  ncvar <- ncvar_def(name     = varid(x)[1], 
                     dim      = list(dimS,dimT), 
                     units    = unique(unit(x)),
                     longname = '5 minutes accumulated precipitation (0,1,2,3,4 minutes)', #tolower(attr(x,'longname')[1]),
                     prec     = "double",
                     verbose  = verbose)
  if (file.exists(file.path(path, filename)) & force) 
    file.remove(file.path(path, filename))
  ncid <- nc_create(filename = file.path(path, filename),
                    vars     = list(ncvar,lonid,xid,latid,yid,altid,locid,stid,cntrid),
                    verbose  = verbose)
  ## Writing the data into the file
  
  start <- c(1,1); 
  count <- c(length(stid(x)),length(index(x)))
  ## var precip
  ncvar_put(nc = ncid, 
            varid = ncvar,
            vals = as.numeric(x),
            start = start,
            count = count)
  
  ## Put variables
  ncvar_put(nc = ncid, varid = lonid, vals = as.numeric(lon(x)), start = 1, count = length(lon(x)))
  ncvar_put(nc = ncid, varid = latid, vals = as.numeric(lat(x)), start = 1, count = length(lat(x)))
  # Convert lon and lat to X and Y
  # browser()
  XY <- LatLon2UTM(lat = lat(x), lon = lon(x), zone = 33)
  ncvar_put(nc = ncid, varid = yid, vals = as.numeric(XY[[1]]), start = 1, count = length(stid(x)))
  ncvar_put(nc = ncid, varid = xid, vals = as.numeric(XY[[2]]), start = 1, count = length(stid(x)))
  ncvar_put(nc = ncid, varid = stid, vals = as.numeric(stid(x)), start = start, count = c(namelength,length(stid(x))))
  ncvar_put(nc = ncid, varid = locid, vals = as.character(loc(x)), start = start, count = c(namelength,length(stid(x))))
  ncvar_put(nc = ncid, varid = cntrid, vals = as.character(cntr(x)), start = start, count = c(namelength,length(stid(x))))
  
  
  ## Put global attributes
  if (verbose) print('Writing global attributes')
  ncatt_put(ncid, 0, 'institution', 'Norwegian Meteorological Institute')
  ncatt_put(ncid, 0, 'date_created', as.character(Sys.time()))
  ncatt_put(ncid, 0, 'Conventions', 'CF-1.6')
  ncatt_put(ncid, 0, 'title', 'Observations from MET Norway weather stations')
  ncatt_put(ncid, 0, 'institution', 'Norwegian Meteorological Institute')
  ncatt_put(ncid, 0, 'source', 'Meterological surface observations via frost.met.no')
  ncatt_put(ncid, 0, 'history', 'frost write netcdf')
  ncatt_put(ncid, 0, 'references', '') 
  ncatt_put(ncid, 0, 'acknowledgment', 'MET Norway via frost.met.no')
  ncatt_put(ncid, 0, 'comment', '5 minutes observations data retrieved from frost.met.no')
  ncatt_put(ncid, 0, 'creator_email', 'abdelakderm@met.no')
  ncatt_put(ncid, 0, 'creator_name', 'Norwegian Meteorological Institute')
  ncatt_put(ncid, 0, 'creator_url', 'https://www.met.no')
  ncatt_put(ncid, 0, 'geospatial_bounds', '')
  ncatt_put(ncid, 0, 'geospatial_bounds_crs', 'latlon')
  ncatt_put(ncid, 0, 'geospatial_lat_max', max(lat(x),na.rm = TRUE))
  ncatt_put(ncid, 0, 'geospatial_lat_min', min(lat(x),na.rm = TRUE))
  ncatt_put(ncid, 0, 'geospatial_lon_max', max(lon(x),na.rm = TRUE))
  ncatt_put(ncid, 0, 'geospatial_lon_min', min(lon(x),na.rm = TRUE))
  ncatt_put(ncid, 0, 'id','add doi here if needed')
  ncatt_put(ncid, 0, 'keywords', 'observations')
  ncatt_put(ncid, 0, 'metadata_link', '')
  ncatt_put(ncid, 0, 'summary', 'Surface meteorological observations from the observation network operated by the Norwegian Meteorological Institute.')
  #ncatt_put(ncid, 0, 'time_coverage_start', it.start)
  #ncatt_put(ncid, 0, 'time_coverage_end', it.end)
  #ncatt_put(ncid, 0, 'featureType', 'timeSeries')
  
  if (verbose) print('Closing connection to NetCDF file')
  nc_close(ncid)
  if (verbose) print('Done!')
}
```

